---
title: "Initial experiment - sensor recording"
author: "João Marreiros & Lisa Schunk"
date: "`r Sys.Date()`"
output: word_document
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```


# Goal of the script
This script reads the CSV files generated with the sensors (SMARTTESTER) during an initaial experiment in order to check the functionality of the sensors.    
The script will:

1. Read in the original CSV files, combines them and organise the data 
2. Plot the data
3. Write an XLSX-file and save an R object ready for further analysis in R 



```{r}
dir_in <- "analysis/raw_data/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(ggplot2)
library(readr)
library(reshape2)
library(ggpubr)
library(magrittr)
library(wesanderson)
library(tools)
library(chron)

```

---

# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                      row.names = NULL)

```


The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in

```


# Read in original data
```{r}
# the data contains information about force, friction, length and velocity

force <- read_table2(data_file[1])
friction <- read_table2(data_file[2])
length <- read_table2(data_file[3])
depth <- read_table2(data_file[4])
velocity <- read_table2(data_file[5])

```


# Organise the data 
```{r}
# fills empty fields

# force
data_force <- as.data.frame(force)
data_force <- melt(data_force, id = "steps")

# friction
data_friction <- as.data.frame(friction)
data_friction <- melt(data_friction, id = "steps")

# length
data_length <- as.data.frame(length)
data_length[is.na(data_length)] <- 0
data_lenght <- melt(data_length, id = "steps")

# depth
data_depth <- as.data.frame(depth)
data_depth <- melt(data_depth, id = "steps")

# velocity
data_velocity <- as.data.frame(velocity)
data_velocity[is.na(data_velocity)] <- 0
data_velocity <- melt(data_velocity, id = "steps")

```

# Plot the data 
```{r}
# force
forceplot <- ggplot(data_force, aes(x = steps, y = value, colour = variable)) +
             geom_line(alpha = 0.3) + 
             theme_classic() +
             labs(x = "distance [mm]", y = "force [N]", 
             title = "") + 
             theme(legend.position = "none") +
             scale_colour_manual(values = wes_palette(n = 50, name =
             "Moonrise2", type = "continuous")) 
           
# friction
frictionplot <- ggplot(data_friction, aes(x = steps, y = value, colour = 
                variable)) + geom_line(alpha = 0.3) +
                theme_classic() +
                labs(x = "distance [mm]", 
                y = "force [N]", title = "", colour = "Strokes") +
                theme(legend.position = "none") +
                scale_colour_manual(values = wes_palette(n = 50, name =
                "Moonrise2", type = "continuous")) 

# velocity
velocityplot <- ggplot(data_velocity, aes(x = steps, y = value, colour =
                variable)) + geom_line(alpha=0.3) +
                theme_classic() +
                geom_line(alpha = 0.3) + 
                labs(x = "distance [mm]", y = "velocity [mm/s]",                                  title = "", colour = "Strokes") + 
                theme(legend.position = "none") +
                scale_colour_manual(values = wes_palette(n = 50, name =
                "Moonrise2", type = "continuous")) 

# depth
depthplot <- ggplot(data_depth, aes(x = steps, y = value, colour = variable)) +
             geom_line(alpha = 0.3) + 
             theme_classic() + 
             labs(x = "distance [mm]", y = "depth [mm]",
             title = "", colour = "Strokes") + theme(legend.position = "none") +
             scale_colour_manual(values = wes_palette(n = 50, name =
             "Moonrise2", type = "continuous")) 

# plot all together
all <- ggarrange(forceplot, frictionplot,  velocityplot, depthplot,
       labels = c("Force", "Friction", "Velocity", "Depth"),
       ncol = 2, nrow = 2) 
       

print(all)

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "all", ".pdf")
ggsave(filename = file_out, plot = all, path = dir_out, device = "pdf")

```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
