---
title: "Volume and weight measurements - Tool function experiment"
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
    html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script reads the xlsx file (two types of measurements: weight measured with a weighing scale and volume calculated based on 3D models.  
The script will:

1. Read in the original xlsx file and organise the data 
2. Plot the data 
3. Write an XLSX-file and save an R object ready for further analysis in R 



```{r}
dir_in <- "analysis_VW/raw_data/"
dir_out <- "analysis_VW/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.


---

# Load packages
```{r Libraries}
pack_to_load <- c("tidyverse", "readr", "openxlsx", "tools", "ggplot2", "readxl",
                  "R.utils","patchwork", "ggsci", "ggfortify", "wesanderson", "doBy")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```


---


# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                  row.names = NULL)

```


# Read in original xlsx-file
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 1, colNames = TRUE, 
                      rowNames = FALSE, skipEmptyRows = FALSE) 
str(imp_data)

# get new order 
imp_data$cycle <- factor(imp_data$cycle, levels = c("before", "after"))
imp_data$edge_angle <- factor(imp_data$edge_angle, levels = c("35", "45"))

flint <- imp_data[1:24,]
lydite <- imp_data[25:48,]

```


# Data analsysis - plot 
## Flint & lydite -  cutting 
```{r}
# flint 
f_cutting_v <- ggplot(data = flint[grep("cutting", flint[["motion"]]), ], 
                       aes_string(x = "cycle", y = "volume_mm3", label = "ID")) +
               geom_point(size = 3, colour = "#046C9A") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#046C9A") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "flint", x = NULL, y = "volume [mm3]") +
               theme_classic()
 
print(f_cutting_v)
 
f_cutting_w <- ggplot(data = flint[grep("cutting", flint[["motion"]]), ], 
                       aes_string(x = "cycle", y = "weight_g", label = "ID")) +
               geom_point(size = 3, colour = "#046C9A") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), , colour = "#046C9A") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "flint", x = NULL, y = "weight [g]") +
               theme_classic()
 
print(f_cutting_w) 

# lydite  
l_cutting_v <- ggplot(data = lydite[grep("cutting", lydite[["motion"]]), ], 
                       aes_string(x = "cycle", y = "volume_mm3", label = "ID")) +
               geom_point(size = 3, colour = "#F2AD00") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#F2AD00") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "lydite", x = NULL, y = "volume [mm3]") +
               theme_classic()
 
print(l_cutting_v)
 
l_cutting_w <- ggplot(data = lydite[grep("cutting", lydite[["motion"]]), ], 
                       aes_string(x = "cycle", y = "weight_g", label = "ID")) +
               geom_point(size = 3, colour = "#F2AD00") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#F2AD00") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "lydite", x = NULL, y = "weight [g]") +
               theme_classic()
 
print(l_cutting_w) 
 

# combine the flint and the lydite plots 
cutting <- f_cutting_v + f_cutting_w + l_cutting_v + l_cutting_w + 
  plot_layout(width = c(2/4, 2/4))  

print(cutting)

#save to PDF
file_out <- paste0(basename(file_path_sans_ext(data_file)), "_cutting.pdf")
ggsave(filename = file_out, plot = cutting, path = dir_out, device = "pdf")
  
```


# Data analsysis - plot 
## Flint & lydite -  carving 
```{r}
# flint
f_carving_v <- ggplot(data = flint[grep("carving", flint[["motion"]]), ], 
                       aes_string(x = "cycle", y = "volume_mm3", label = "ID")) +
               geom_point(size = 3, colour = "#046C9A") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#046C9A") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "flint", x = NULL, y = "volume [mm3]") +
               theme_classic()

print(f_carving_v)

f_carving_w <- ggplot(data = flint[grep("carving", flint[["motion"]]), ], 
                       aes_string(x = "cycle", y = "weight_g", label = "ID")) +
               geom_point(size = 3, colour = "#046C9A") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), , colour = "#046C9A") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "flint", x = NULL, y = "weight [g]") +
               theme_classic()
 
print(f_carving_w) 

 
# lydite  
l_carving_v <- ggplot(data = lydite[grep("carving", lydite[["motion"]]), ], 
                       aes_string(x = "cycle", y = "volume_mm3", label = "ID")) +
               geom_point(size = 3, colour = "#F2AD00") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#F2AD00") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "lydite", x = NULL, y = "volume [mm3]") +
               theme_classic()
 
print(l_carving_v)
 
l_carving_w <- ggplot(data = lydite[grep("carving", lydite[["motion"]]), ], 
                       aes_string(x = "cycle", y = "weight_g", label = "ID")) +
               geom_point(size = 3, colour = "#F2AD00") + 
               geom_text(size = 3, hjust = 0.5, vjust = 1.7) +
               geom_line(aes(group = ID), colour = "#F2AD00") + 
               theme(legend.position = "none") +
               facet_wrap(edge_angle ~ motion) + 
               labs(title = "lydite", x = NULL, y = "weight [g]") +
               theme_classic()
 
print(l_carving_w) 
 
 
# combine the flint and the lydite plots 
carving <- f_carving_v + f_carving_w + l_carving_v + l_carving_w + plot_layout(width = c(2/4, 2/4), guides = 'collect')  

print(carving)

#save to PDF
file_out <- paste0(basename(file_path_sans_ext(data_file)), "_carving.pdf")
ggsave(filename = file_out, plot = carving, path = dir_out, device = "pdf")

```

The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("RStudioVersion.txt", n = 1)`.



# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```

--- 

END OF SCRIPT