---
title: "Plots - cut marks/groovres"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
    html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots the variables to visualise the results of the tool function experiment.  


```{r}
dir_in <- "analysis_S.wide/derived_data/"
dir_out <- "analysis_S.wide/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---

# Load packages
```{r Libraries}
pack_to_load <- c("tidyverse", "R.utils", "ggplot2", "tools", "patchwork", "ggsci", "ggfortify", "wesanderson", "doBy")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```



---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)

```


The imported file is:  
```{r, echo = FALSE}
data_file

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 19:length(imp_data)

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep = "")

```


---

## Convert nm to µm for MaxDepth and add a new column for type ("Data point" or "Mean")
```{r}
imp_data <- mutate(imp_data, Max.depth.µm = Maximum.depth/1000, Type = "Data point")
imp_data

```

---


# Plot the data: width against depth  
## cutting (BP I.)
```{r}
# filter the data for the cutting only 
cutting <- filter(imp_data, Movement == "cutting")

BP.I <- ggplot(cutting, aes(x = Width, y = Depth, label = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   geom_text(size = 3, hjust = 0.5, vjust = 1.5) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(x = "widht [µm]", y = "depth [µm]", title = "cutting") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(BP.I)


#save to PDF
file_out <- paste0(basename(file_path_sans_ext(data_file)), "_cutting.pdf")
ggsave(filename = file_out, plot = BP.I, path = dir_out, device = "pdf")

```


## carving (BP II. & III.)
```{r}
# filter the data for the carving only 
carving <- filter(imp_data, Movement == "carving")
# remove "outlier" (FLT8-10)
carving <- carving[-4, ]

BP.II_III <- ggplot(carving, aes(x = Width, y = Depth, label = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   geom_text(size = 3, hjust = 0.5, vjust = 2.5) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(x = "widht [µm]", y = "depth [µm]", title = "carving") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(BP.II_III)


#save to PDF
file_out <- paste0(basename(file_path_sans_ext(data_file)), "_carving.pdf")
ggsave(filename = file_out, plot = BP.II_III, path = dir_out, device = "pdf")

```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("RStudioVersion.txt", n = 1)`.


---

# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```


---

END OF SCRIPT
