---
title: "Plots - cut marks/groovres"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output: word_document
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots the variables to visualise the results of the tool function experiment.  



```{r}
dir_in <- "analysis_S.wide/derived_data/"
dir_out <- "analysis_S.wide/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(ggsci)
library(ggfortify)
library(wesanderson)
library(doBy)
library(ggfortify)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 19:length(imp_data)

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```


---

## Convert nm to µm for MaxDepth and add a new column for type ("Data point" or "Mean")
```{r}
imp_data <- mutate(imp_data, Max.depth.µm = Maximum.depth.1/1000, Max.depth.µm = Maximum.depth.1/1000)
imp_data

# converts Sample from character into factor 
imp_data[["Sample"]] <- factor(imp_data[["Sample"]])

```

---

# Plot the maximum depth (facet plot = 1 plot per bone plate)
## bone plate I. (cutting)
```{r}

TFE <- filter(imp_data, Experiment == "TFE")
TFE.I <- filter(TFE, Bone.plate == "BP-I")

# orders Sample differently 
TFE.I[["Sample"]] <- ordered(TFE.I[["Sample"]], levels = c("FLT8-4", "FLT8-5", "FLT8-6", "LYDIT5-5", "LYDIT5-6", "LYDIT5-7", "FLT8-1", "FLT8-2", "FLT8-3", "LYDIT5-2", "LYDIT5-3", "LYDIT5-4"))


p.TFE.I <- ggplot(TFE.I, aes(y = Max.depth.µm, x = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(y = "max. depth [µm]", x = "sample", title = "cutting") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.I)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.I", ".pdf")
ggsave(filename = file_out, plot = p.TFE.I, path = dir_out, device = "pdf")


```


## bone plate II. + III. (carving)
```{r}

TFE <- filter(imp_data, Experiment == "TFE")
TFE.II <- filter(TFE, Bone.plate == "BP-II" | Bone.plate == "BP-III")

# removes a additional test samples used for scraping  
TFE.II <- TFE.II[1:12, ]

# orders Sample differently 
TFE.II[["Sample"]] <- ordered(TFE.II[["Sample"]], levels = c("FLT8-7", "FLT8-8", "FLT8-9", "LYDIT5-11", "LYDIT5-12", "LYDIT5-13", "FLT8-10", "FLT8-11", "FLT8-12", "LYDIT5-8", "LYDIT5-9", "LYDIT5-10"))


p.TFE.II <- ggplot(TFE.II, aes(y = Max.depth.µm, x = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(y = "max. depth [µm]", x = "sample", title = "carving") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.II)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.II", ".pdf")
ggsave(filename = file_out, plot = p.TFE.II, path = dir_out, device = "pdf")


```


## bone plate II. + III. (carving)
### without sample FLT8-10 ("outlier")
```{r}

# removes FLT8-10  
TFE.IIb <- TFE.II[-4, ]


p.TFE.IIb <- ggplot(TFE.IIb, aes(y = Max.depth.µm, x = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(y = "max. depth [µm]", x = "sample", title = "carving") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.IIb)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.IIb", ".pdf")
ggsave(filename = file_out, plot = p.TFE.IIb, path = dir_out, device = "pdf")


```


# Plot the width against the depth (facet plot = 1 plot per bone plate)
## bone plate I. (cutting)
```{r}

TFE <- filter(imp_data, Experiment == "TFE")
TFE.I <- filter(TFE, Bone.plate == "BP-I")


p.TFE.I2 <- ggplot(TFE.I, aes(x = Distance.1, y = Distance.2, label = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   geom_text(size = 3, hjust = 0.5, vjust = 1.5) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(x = "widht [µm]", y = "depth [µm]", title = "cutting") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.I2)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.I2", ".pdf")
ggsave(filename = file_out, plot = p.TFE.I2, path = dir_out, device = "pdf")


```


## bone plate II. + III. (carving)
```{r}

TFE <- filter(imp_data, Experiment == "TFE")
TFE.II <- filter(TFE, Bone.plate == "BP-II" | Bone.plate == "BP-III")

TFE.II <- TFE.II[1:12, ]

p.TFE.II2 <- ggplot(TFE.II, aes(x = Distance.1, y = Distance.2, label = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   geom_text(size = 3, hjust = 0.5, vjust = 2.5) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(x = "widht [µm]", y = "depth [µm]", title = "carving") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.II2)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.II2", ".pdf")
ggsave(filename = file_out, plot = p.TFE.II2, path = dir_out, device = "pdf")


```


## bone plate II. + III. (carving)
### without sample FLT8-10 ("outlier")
```{r}

p.TFE.II2b <- ggplot(TFE.IIb, aes(x = Distance.1, y = Distance.2, label = Sample)) +
                   geom_point(aes(colour = Raw.material, shape = Edge.angle), size = 4) +
                   geom_text(size = 3, hjust = 0.5, vjust = 2.5) +
                   scale_y_reverse() +
                   theme_classic() +
                   guides(colour = "none") +
                   labs(x = "widht [µm]", y = "depth [µm]", title = "carving") +
                   labs(shape = "edge angle") +
                   scale_colour_manual(values = c("#046C9A", "#F2AD00"))
                 
plot(p.TFE.II2b)


#save to PDF
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_TFE.II2b", ".pdf")
ggsave(filename = file_out, plot = p.TFE.II2b, path = dir_out, device = "pdf")


```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
